Efficient Handling of String-Number Conversion

### 2019-11-21 draft.tex
some comments to section 5

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\hide}[1]{}
\newcommand{\tool}{{\textsf{Z3-PFA}}}

\newcommand{\nat}{\mathbb{N}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\todo}[1]{{\color{blue}TODO: #1}}
\newcommand{\lh}[1]{{\color{orange}Lukas: #1}}
\newcommand{\changed}[1]{{\color{orange}#1}}
\newcommand{\yfc}[1]{{\color{blue}YFC: #1}}
\newcommand{\petr}[1]{{\color{pink}Petr: #1}}
\newcommand{\chatAt}[2]{\mbox{\textsf{charAt}($#1$, $#2$)}}
\newcommand{\ite}[3]{\mbox{\textsf{ite}($#1$, $#2$, $#3$)}}
\newcommand{\sti}[1]{\mbox{\textsf{toNum}($#1$)}}
\newcommand{\its}[1]{\mbox{\textsf{toStr}($#1$)}}
\newcommand{\varn}{\mbox{$\mathbb{V}_{\mathbb{Z}}$}}
\newcommand{\vars}{\mbox{$\mathbb{V}_{\Sigma^*}$}}
\newcommand{\cvars}{\mbox{$\mathbb{V}_{\Sigma_\epsilon}$}}
\newcommand{\pvars}{\mbox{$\mathbb{V}_{\sharp}$}}
%\newcommand{\cvar}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvarone}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvartwo}{\mbox{$V'_{\Sigma_\epsilon}$}}
\newcommand{\cvarone}{V}
\newcommand{\cvartwo}{V'}
\newcommand{\cvar}{V}
\newcommand{\modelsof}[1]{[\![#1]\!]}
\newcommand{\true}{\mbox{$\mathsf{true}$}}
\newcommand{\false}{\mbox{$\mathsf{false}$}}
\newcommand{\enc}[1]{[\![#1]\!]}
%\newcommand{\parikhof}[1]{\mathit{Parikh}{(#1)}}
\newcommand{\parikhof}[1]{\mathbb{P}{(#1)}}
\newcommand{\parikhwof}[2]{|#1|_{#2}}
%\newcommand{\semof}[1]{||#1||}
\newcommand{\semof}[1]{\modelsof{#1}}
\newcommand{\decode}[1]{\mathit{decode_{#1}}}
%\newcommand{\parikhfof}[1]{\Phi_{\parikhof{#1}}}
\newcommand{\parikhfof}[1]{\Phi_{\mathbb{P}}(#1)} %shouldn't it be something that takes automaton as parameter and returns formula?
\newcommand{\pim}{I_{\#}}
%\newcommand{\syncop}{\Cap}
\newcommand{\syncop}{\times}
\newcommand{\syncof}[2]{#1 \syncop #2}
\newcommand{\syncfof}[2]{\Psi_{\syncof {#1} {#2}}}
\newcommand{\syncT}{T_\syncop}
\newcommand{\pvarsof}[1]{\#{#1}}
\newcommand{\pvar}{\pvarsof V}
\newcommand{\pvarone}{\pvarsof V}
\newcommand{\pvartwo}{\pvarsof {V'}}
\newcommand{\defeq}{::=}
\newcommand{\encode}[1]{\mathit{encode}_{#1}}
\newcommand{\eqwrt}[1]{=_{#1}}
\newcommand{\iequiv}{\equiv}
\newcommand{\restrict}[2]{#1_{#2}}
\newcommand{\underf}[2]{\mathit{under}_#1(#2)}
\newcommand{\paf}{\psi}

@@

The domain restriction is formally defined by restricting the domain of each string variable by a chosen PFA. 
%
Namely, assuming that $X$ is the set of string variables of $\insc$,
a \emph{flat domain restriction} for $\insc$ is a mapping $R$ that assigns to every variable $x\in X$ a PFA $R(x)$ \changed{over character variables $V_x$.}
string %
We require that these PFA operate over pairwise disjoint sets of character variables\changed{, that is if $x\neq y$ then $V_x \cap V_y = \emptyset$.}
%
The particular choice of a PFA 
for each variable  depends on the strategy used in the implementation, and will be discussed in Section~\ref{section:implementation}.
%
The flattening $\outlf$ will be  built inductively to the structure of $\insc$.
Namely, in the following subsections we show how to build a flattening $\underf R {\phi}$ for every atomic string constraint $\phi$, 
and for a conjunction of string constraints, we let $\underf R {\phi \land \phi'} \defeq \underf R {\phi} \land \underf R {\phi'}$.
We then let $\outlf \defeq \underf R \insc$

The semantics of a string constraint $\phi$ restricted by $R$ is then defined as 
$\semof{\phi}^R = \{I\in\semof \phi \mid \forall x\in X:I(x)\in\semof{R(x)}\}$. 
The following sections will be devoted to describing the flattening of the atomic string constraints. 
The correctness of the entire construction of $\outlf$ is expressed by  Theorem~\ref{theorem:correct}:
\begin{theorem}\label{theorem:correct}
$\decode R (\semof {\outlf}) = \semof{\insc}_R$ 
\end{theorem}

The theorem can be proved by induction to the structure of $\insc$,
however, for the induction step to go through, we will need a more meticulous view into the correspondence of $\phi$ and $\underf R \phi$ than just equality of their semantics. 
Particularly, we will need to
 ensure that $\underf R \phi$ captures exactly all $R$-encodings of $\semof \phi^R$
(indeed, notice that if it would be allowed to capture only some of the encodings, then for instance $\underf R {\phi} \land \underf R {\phi'}$ could only underapproximate $\semof{\phi\land\phi'}^R$).

To formalise this correctness argument, we define an additional notation.
For an interpretation $I$ of string variables, 
we define its $R$-encoding as the set $\encode R(I) = \bigcup_{x\in X} \encode \pa(I(x))$ of interpretations of $V_R\cup \#V_R$, containing combinations of interpretations of character variables of each $R(x)$ that encode the $I$.
%
\changed{
Further, we define an $R$-encoding for an interpretation $I \cup I'$ of a general string constraint, consiting of an interpretation $I$ of string variables and an interpretation $I'$ of integer variables.
%
An $R$-encoding of $I \cup I'$ is a union  $e \cup I'$ of an encoding $e\in \encode R(I)$ and the original integer interpretation $I'$ which satisfies that for every string leght variable $|x|$ in the domain of $I'$,
$|x| = \sum_{v\in V_x,I(v)\neq\epsilon}\#v$. 
That is, the interpretation of string variables is encoded as usual, the interpretation of integer variables is kept as is, 
and for each string legth variable $|x|$ in the domain of $I'$, 
the value of $|x|$ is consistent with the encoding of $x$ by variables $V_x$ and $\#V_x$.
The set of all $R$-encodings of $(I\cup I')$ is then denoted $\encode R {(I\cup I')}$}
%
%Further, if $I \cup I'$ is an interpretation of a general string constraint, that is, an interpretation $I$ of string variables $X$ united with an interpretation $I'$ of integer variables $N$,
%%
%then an $R$-encoding of $I \cup I'$ is any interpretation $e \cup I$ with  $e:V_R \cup \#V_R \cup \Simga_\epsilon$ 
%such that $\encode R {(I\cup I')} \defeq \{e \cup I' \mid e\in \encode R(I)\}$ of interpretations of variables $V_R \cup \#V_R \cup N$. That is, the interpretation of string variables is encoded and the interpretation of integer variables is kept as is.
%
%
and we lift this notation to sets of interpretation in the standard manner, i.e., $\encode R (J) = \bigcup_{I \in J}\encode R (I)$ for a set of interpretations $J$.
%
The inductive argument needed in the proof of the correctness of the under-approximation then reads as
$$\semof{\underf R \phi}_{V_R\cup V_{\#R}}  = \encode R{\semof{\phi}}$$ 
%
In the next sections, 
we will formulate the corresponding correctness lemma for flattenings constructed from each type of atomic string constraints.
%
We note that the restriction of the semantics of $\underf R \phi$ here is needed since $\underf R \phi$ will be constructed with some auxiliary variables.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\hide}[1]{}
\newcommand{\tool}{{\textsf{Z3-PFA}}}
\newcommand{\dr}{\mathcal{R}}
\newcommand{\nat}{\mathbb{N}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\todo}[1]{{\color{blue}TODO: #1}}
\newcommand{\lh}[1]{{\color{orange}Lukas: #1}}
\newcommand{\changed}[1]{{\color{orange}#1}}
\newcommand{\yfc}[1]{{\color{blue}YFC: #1}}
\newcommand{\petr}[1]{{\color{pink}Petr: #1}}
\newcommand{\chatAt}[2]{\mbox{\textsf{charAt}($#1$, $#2$)}}
\newcommand{\ite}[3]{\mbox{\textsf{ite}($#1$, $#2$, $#3$)}}
\newcommand{\sti}[1]{\mbox{\textsf{toNum}($#1$)}}
\newcommand{\its}[1]{\mbox{\textsf{toStr}($#1$)}}
\newcommand{\varn}{\mbox{$\mathbb{V}_{\mathbb{Z}}$}}
\newcommand{\vars}{\mbox{$\mathbb{V}_{\Sigma^*}$}}
\newcommand{\cvars}{\mbox{$\mathbb{V}_{\Sigma_\epsilon}$}}
\newcommand{\pvars}{\mbox{$\mathbb{V}_{\sharp}$}}
%\newcommand{\cvar}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvarone}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvartwo}{\mbox{$V'_{\Sigma_\epsilon}$}}
\newcommand{\cvarone}{V}
\newcommand{\cvartwo}{V'}
\newcommand{\cvar}{V}
\newcommand{\modelsof}[1]{[\![#1]\!]}
\newcommand{\true}{\mbox{$\mathsf{true}$}}
\newcommand{\false}{\mbox{$\mathsf{false}$}}
\newcommand{\enc}[1]{[\![#1]\!]}
%\newcommand{\parikhof}[1]{\mathit{Parikh}{(#1)}}
\newcommand{\parikhof}[1]{\mathbb{P}{(#1)}}
\newcommand{\parikhwof}[2]{|#1|_{#2}}
%\newcommand{\semof}[1]{||#1||}
\newcommand{\semof}[1]{\modelsof{#1}}
\newcommand{\decode}[1]{\mathit{decode_{#1}}}
%\newcommand{\parikhfof}[1]{\Phi_{\parikhof{#1}}}
\newcommand{\parikhfof}[1]{\Phi_{\mathbb{P}}(#1)} %shouldn't it be something that takes automaton as parameter and returns formula?
\newcommand{\pim}{I_{\#}}
%\newcommand{\syncop}{\Cap}
\newcommand{\syncop}{\times}
\newcommand{\syncof}[2]{#1 \syncop #2}
\newcommand{\syncfof}[2]{\Psi_{\syncof {#1} {#2}}}
\newcommand{\syncT}{T_\syncop}
\newcommand{\pvarsof}[1]{\#{#1}}
\newcommand{\pvar}{\pvarsof V}
\newcommand{\pvarone}{\pvarsof V}
\newcommand{\pvartwo}{\pvarsof {V'}}
\newcommand{\defeq}{::=}
\newcommand{\encode}[1]{\mathit{encode}_{#1}}
\newcommand{\eqwrt}[1]{=_{#1}}
\newcommand{\iequiv}{\equiv}
\newcommand{\restrict}[2]{#1_{#2}}
\newcommand{\underf}[2]{\mathit{under}_#1(#2)}
\newcommand{\paf}{\psi}

@@

The domain restriction is formally defined by restricting the domain of each string variable by a chosen PFA. 
%
Namely, assuming that $X$ is the set of string variables of $\insc$,
a \emph{flat domain restriction} for $\insc$ is a mapping $\dr$ that assigns to every variable $x\in X$ a PFA $\dr(x)$ over character variables $V_x$.
We require that these PFA operate over pairwise disjoint sets of character variables, that is if $x\neq y$ then $V_x \cap V_y = \emptyset$.
%
The particular choice of a PFA 
for each variable  depends on the strategy used in the implementation, and will be discussed in Section~\ref{section:evaluation}.
%
The flattening $\outlf$ will be  built inductively to the structure of $\insc$.
Namely, in the following subsections we show how to build a flattening $\underf \dr {\phi}$ for every atomic string constraint $\phi$, 
and for a conjunction of string constraints, we let $\underf \dr {\phi \land \phi'} \defeq \underf \dr {\phi} \land \underf \dr {\phi'}$.
We then let $\outlf \defeq \underf \dr \insc$

The semantics of a string constraint $\phi$ restricted by $\dr$ is then defined as 
$\semof{\phi}^\dr = \{I\in\semof \phi \mid \forall x\in X:I(x)\in\semof{\dr(x)}\} \subseteq \vars\times \varn \rightarrow \Sigma^* \times\nat$.
The following sections will be devoted to describing the flattening of the atomic string constraints. 
The correctness of the entire construction of $\outlf$ is expressed by  Theorem~\ref{theorem:correct}:
\begin{theorem}\label{theorem:correct}
$\decode R (\semof {\outlf}) = \semof{\insc}^\dr$ 
\end{theorem}

The theorem can be proved by a structural induction over $\insc$.
However, for the induction step to go through, we will need a more meticulous view into the correspondence of $\phi$ and $\underf \dr \phi$ than just equality of their semantics. \yfc{I do not understand the previous sentence}
Particularly, we will need to
 ensure that $\underf \dr \phi$ captures exactly all $\dr$-encodings of $\semof \phi^\dr$
(indeed, notice that if it would be allowed to capture only some of the encodings, then for instance $\underf \dr {\phi} \land \underf \dr {\phi'}$ could only underapproximate $\semof{\phi\land\phi'}^\dr$).

To formalize this correctness argument, we define an additional notation.
For an interpretation $I$ of string variables, 
we define its $\dr$-encoding as the set $\encode \dr(I) = \bigcup_{x\in X} \encode {\dr(x)} (I(x))$ of interpretations of $V_\dr\cup \#V_\dr$, containing combinations of interpretations of character variables of each $\dr(x)$ that encode the $I$. \yfc{the sentence is difficult to parse. I did not find the definition of $V_\dr\cup \#V_\dr$.}
%
\changed{
Further, we define an $\dr$-encoding for an interpretation $I \cup I'$ of a general string constraint, consisting of an interpretation $I$ of string variables and an interpretation $I'$ of integer variables.
%
An $\dr$-encoding of $I \cup I'$ is a union  $e \cup I'$ of an encoding $e\in \encode \dr(I)$ and the original integer interpretation $I'$ which satisfies that for every string length variable $|x|$ in the domain of $I'$,
$|x| = \sum_{v\in V_x,I(v)\neq\epsilon}\#v$. 
That is, the interpretation of string variables is encoded as usual, the interpretation of integer variables is kept as is, 
and for each string length variable $|x|$ in the domain of $I'$, 
the value of $|x|$ is consistent with the encoding of $x$ by variables $V_x$ and $\#V_x$.
The set of all $\dr$-encodings of $(I\cup I')$ is then denoted $\encode \dr {(I\cup I')}$}\yfc{need a serious rewrite :-) Maybe I need to know why you define so many encodings...}
%
%Further, if $I \cup I'$ is an interpretation of a general string constraint, that is, an interpretation $I$ of string variables $X$ united with an interpretation $I'$ of integer variables $N$,
%%
%then an $R$-encoding of $I \cup I'$ is any interpretation $e \cup I$ with  $e:V_R \cup \#V_R \cup \Simga_\epsilon$ 
%such that $\encode R {(I\cup I')} \defeq \{e \cup I' \mid e\in \encode R(I)\}$ of interpretations of variables $V_R \cup \#V_R \cup N$. That is, the interpretation of string variables is encoded and the interpretation of integer variables is kept as is.
%
%
and we lift this notation to sets of interpretation in the standard manner, i.e., $\encode \dr (J) = \bigcup_{I \in J}\encode \dr (I)$ for a set of interpretations $J$.
%
The inductive argument needed in the correctness proof of the under-approximation then reads as
$$\semof{\underf \dr \phi}_{V_\dr\cup V_{\#\dr}}  = \encode \dr{\semof{\phi}}$$ 
%
In the next sections, 
we will formulate the corresponding correctness lemma for flattening constructed from each type of atomic string constraints.
%
We note that the restriction of the semantics of $\underf \dr \phi$ here is needed since $\underf \dr \phi$ will be constructed with some auxiliary variables.
